---
title: "Fraxinus_americana"
author: "Lauren Jenkins"
date: "12/7/2021"
output: html_document
---

111 = full species (fraxinus_americana)
222 = species abreviated (Frax_Amer)

Setup environment:
```{r, include = FALSE}
rm(list = ls())
library(dplyr)
library(ggplot2)

library(raster)
library(rgdal)
library(enmSdm)
library(terra)
library(maxnet)
library(sf)

library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)

setwd('/Volumes/LJ MacBook Backup/MOBOT/PVMvsENM')
load('./PCA_Beyer')
load('./03 - Modeling Workspace - Frax_Amer Cleaning')
```

Constants:
```{r}
ll <- c('longitude', 'latitude')
species <- 'fraxinus_americana'
speciesAb <- 'Frax_Amer'
```

Load environmental PCA data:
```{r, include = FALSE}
pcPrediction <- list()
for (i in 1:length(clim)) {
  names(clim[[i]]) <- c("bio1", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16",
                        "bio17", "bio18", "bio19", "bio4", "bio5", "bio6", "bio7",
                        "bio8", "bio9", "cloudiness", "relative_humidity")
  pcPrediction[i] <- raster::predict(clim[[i]], pca, index = 1:3)
  names(pcPrediction[[i]]) <- paste0("pc", 1:3, "_", (i-1)*1000, "KYBP")
}
```

```{r}
envDataPca <- stack(pcPrediction)
# plot(envDataPca)
envPresences <- stack(envDataPca[[1:3]])
names(envPresences) <- paste0('pca', 1:3)
```

Define study region & extent.
```{r}
if (file.exists('./data_and_analyses/study_region/Study Region.Rdata') & 
    file.exists('./data_and_analyses/study_region/Study Region Extent.Rdata')) {
  load('./data_and_analyses/study_region/Study Region.Rdata')
  load('./data_and_analyses/study_region/Study Region Extent.Rdata')
} else {
  studyRegion <- rgdal::readOGR(dsn = './data_and_analyses/study_region', 'study_region')
  studyRegion <- crop(studyRegion, extent(-178.2166, 83.7759, -55.90223, 83.6236))
  projection(studyRegion) <- getCRS("WGS84")
  studyExtent <- extent(studyRegion)
  save(studyRegion, file='./data_and_analyses/study_region/Study Region.Rdata', compress=TRUE)
  save(studyExtent, file='./data_and_analyses/study_region/Study Region Extent.Rdata', compress=TRUE)
}
```

Clip environmental PCA's to study extent & visualize:
```{r, include = FALSE}
fileName <- paste0('./data_and_analyses/env_data/Beyer/envDataClipped_', species, '.tif')
if (file.exists(fileName)) {
  envData <- brick(fileName)
  names(envData) <- paste0('pca', 1:3)
} else {
  envDataClipped <- list()
  for (i in 1:nlayers(envPresences)) {
    x <- envPresences[[i]]
    x <- crop(x, studyExtent)
    x <- mask(x, studyRegion)
    projection(x) <- getCRS("WGS84")
    envDataClipped[[i]] <- x
    envData <- stack(envDataClipped)
    # plot(envData)
    writeRaster(envData, fileName, 
                format = 'GTiff', overwrite = T)
  }
}
```

Load records for given species.
```{r}
records <- paste0('./cleaning_records/', 
                         species, '_finalRecords.rData')
# records <- paste0('./species_records/02_', species, '_thinned_records.rData')
load(records)
records <- speciesSf_filtered_final # 515 observations
records <- as(records, 'Spatial')
records <- as.data.frame(records)
colnames(records)[26:27] <- c('longitude', 'latitude')
```

Match environmental data to records.
```{r}
envSpecies <- extract(envData, cbind(records$longitude, records$latitude))
envSpecies <- as.data.frame(envSpecies)
records <- cbind(records, envSpecies)
```

Remove records that fall in the water & visualize.
```{r}
# records in the water:
if (any(is.na(rowSums(envSpecies)))) {
  water <- records[which(is.na(rowSums(envSpecies))), ] 
  water <- SpatialPointsDataFrame(water[,ll], data = water,
                                proj4 = getCRS('wgs84', TRUE))
}

# remove records in water from dataset:
if (any(is.na(rowSums(envSpecies)))) records <-
  records[-which(is.na(rowSums(envSpecies))), ]

# visualize the points that fall in the water
# convert to sp object
recordsSp <- SpatialPointsDataFrame(records[, ll], data = records,
                                       proj4 = getCRS('wgs84', TRUE))

# visualize
plot(recordsSp, pch = 16, cex = 0.5, col = "red", 
     main = 'BIEN occurrences thinned')
if (exists("water")) {
  plot(water, col = 'blue', add = TRUE)
}
map("state", add = TRUE)
map("world", add = TRUE)
```

Save workspace.
```{r}
save.image(paste0('./04 - Modeling Workspace - Clipping Environmental Data - ', species))
# load('./04 - Modeling Workspace - Clipping Environmental Data')
```

```{r}
climate <- envData
# generate 10,000 background sites
randomBgSites <- randomPoints(climate, 10000)

# extract environment at sites
randomBgEnv <- extract(climate, randomBgSites)
randomBgEnv <- as.data.frame(randomBgEnv)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv))
if (any(isNa)) {
    randomBgSites <- randomBgSites[-which(isNa), ]
    randomBgEnv <- randomBgEnv[-which(isNa), ]
}

# combine with coordinates and rename coordinate fields
randomBg <- cbind(randomBgSites, randomBgEnv)
names(randomBg)[1:2] <- c('longitude', 'latitude')
dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
save(randomBg, file=
    './Background Sites/Random Background Sites across Study Region.Rdata',
    compress=TRUE)
```

Generate test background sites.
```{r}
load(paste0('./03 - Modeling Workspace - ', speciesAb, ' Cleaning'))

range_buffer_final <- as(range_buffer_final, Class = 'Spatial')
calibRegionSpAlb <- range_buffer_final
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
bgTestSpAlb <- spsample(calibRegionSpAlb, n = 11000, type = 'random', iter = 10)
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
bgTest <- as.data.frame(coordinates(bgTestSp))
names(bgTest) <- ll
calibRegionsSp <- calibRegionSpAlb

rasts <- envData
areas <- area(rasts[[1]])

# extract
thisEnv <- raster::extract(rasts, bgTest)
thisEnv <- as.data.frame(thisEnv)

thisArea <- raster::extract(areas, bgTest)
thisArea <- as.data.frame(thisArea)
names(thisArea) <- 'cellArea_km2'

thisEnv <- cbind(thisArea, thisEnv)
env <- thisEnv

bgTest <- cbind(bgTest, env)
bgTest <- bgTest[complete.cases(bgTest), ]
bgTest <- bgTest[1:min(nrow(bgTest), 10000), ]

bgTestSp <- SpatialPoints(bgTest[ , ll], getCRS('wgs84', TRUE))
bgTestSpAlb <- sp::spTransform(bgTestSp, getCRS('albersNA', TRUE))
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE)
Also, type = cclog (to make the output between 0 & 1)
```{r}
predictors <- c('pca1', 'pca2', 'pca3')
trainData <- rbind(records[ ,predictors], randomBg[ ,predictors])

presBg <- c(rep(1, nrow(records)), rep(0, nrow(randomBg)))

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb, '_Maxent')
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel <- maxnet(p = presBg, data = trainData)
save(envModel, file='./Models/Frax_Amer_Maxent/Model.Rdata', compress=TRUE)
```

```{r}
# write current raster
envMap <- predict(
  climate[[predictors]],
  envModel,
  filename='./Models/Frax_Amer_Maxent/maxentPredictionBeyer0KYBP', 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp <- rasterToPolygons(envMap)
envMapSf <- st_as_sf(envMapSp,
                     coords = c(x = 'longitude', y = 'latitude'),
                      crs = getCRS("wgs84", TRUE))
rangeMapSf <- st_transform(st_as_sf(x = range), getCRS('wgs84', TRUE))
recordsSf <- speciesSf <- st_as_sf(x = records,
                      coords = c(x = 'longitude',
                                 y = 'latitude'),
                      crs = getCRS('wgs84', TRUE))

world <- ne_countries(scale = "medium", returnclass = "sf")
states <- st_as_sf(ne_states(country = 'united states of america'))
canada <- st_as_sf(ne_states(country = 'canada'))

# ggplot(data = world) +
#   theme_bw() +
#   geom_sf(fill = "white") +
#   geom_sf(data = states, fill = NA) + 
#   geom_sf(data = canada, fill = NA) +
#   geom_sf(data = envMapSf) +
#   geom_sf(data = rangeMapSf, 
#           color = 'darkgreen',
#           fill = 'green', 
#           alpha = 0.4, 
#           inherit.aes = FALSE) +
#   geom_sf(data = recordsSf,
#           color = 'red',
#           pch = 16,
#           cex = 0.4) + 
#   coord_sf(xlim = c(min(records$longitude), max(records$longitude)), 
#            ylim = c(min(records$latitude), max(records$latitude)), expand = TRUE) +
#   xlab("Longitude") + 
#   ylab("Latitude") +
#   theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
#                                         size = 0.5), panel.background = element_rect(fill = "lavender"))

```

```{r}
plot(littleRange_FraxAmer, border = 'blue', main = "Maxent Output, Frax_amer")
plot(envMap, add = TRUE)
plot(littleRange_FraxAmer, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

```{r}
save.image('./05 - Modeling Workspace - Frax_amer Model Output')
load('./05 - Modeling Workspace - Frax_amer Model Output')
```


<!-- May want to make a MESS map (Elith et al 2010) to show places projected to the past -->
<!-- In the dismo package -->
