---
title: "Finding perfect model"
author: "Lauren Jenkins"
date: "20 December 2021"
output: html_document
---

Just playing with buffer sizes & pca's

Setup environment:
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(tidyr)

library(raster)
library(rgdal)
library(enmSdm)
library(terra)
library(maxnet)
library(sf)

library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)

setwd('/Volumes/LJ MacBook Backup/MOBOT/PVMvsENM')
```

Load PCA environment and set constants:
```{r}
# load PCA data
load('./PCA_Beyer')

# set constants
ll <- c('longitude', 'latitude')
sp <- 'Fraxinus pennsylvanica'
species <- 'fraxinus_pennsylvanica'
speciesAb <- 'FraxPenn'
speciesAb_ <- 'Frax_Penn'
```

Load records for given species.
```{r}
recordsRaw <- paste0('./cleaning_records/', species, '_finalRecords.rData')
load(recordsRaw)
recordsRaw <- data.frame(speciesSf_filtered_final) # 542 observations
recordsRaw$geometry <- gsub("[c()]", "", recordsRaw$geometry)
recordsRaw <- separate(data = recordsRaw, 
                    col = 'geometry', 
                    into = ll, 
                    sep = "\\,")
recordsRaw$longitude <- as.double(recordsRaw$longitude)
recordsRaw$latitude <- as.double(recordsRaw$latitude)
```

Load cleaned records data.
```{r}
load(paste0('./03 - Modeling Workspace - ', speciesAb_, ' Cleaning'))
rangeMap <- littleRange_FraxPenn
```

Load environmental PCA rasters. If already clipped, load that. Otherwise, clip the data to present.
PC1 = 44.44%
PC2 = 71.67%
PC3 = 81.05%
PC4 = 87.65%
PC5 = 91.6%
PC6 = 96.39%
```{r}
pcPrediction <- list()
# label each env layer by variable and year
for (i in 1:length(clim)) {
  names(clim[[i]]) <- c("bio1", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16",
                        "bio17", "bio18", "bio19", "bio4", "bio5", "bio6", "bio7",
                        "bio8", "bio9", "cloudiness", "relative_humidity")
  pcPrediction[i] <- raster::predict(clim[[i]], pca, index = 1:4)
  names(pcPrediction[[i]]) <- paste0("pc", 1:4, "_", (i-1)*1000, "KYBP")
}

envDataPca <- stack(pcPrediction)

# keep only rasters (first three rasters) from present (0 KYBP)
envPresences <- stack(envDataPca[[1:4]])
names(envPresences) <- paste0('pca', 1:4)

# check projections = wgs84
print("Ensure that the projection of these rasters is WGS84:")
print(paste0("Projection of envPresences = ", projection(envPresences)))

# define study region & extent, load if already defined
if (file.exists('./data_and_analyses/study_region/Study Region.Rdata') & 
    file.exists('./data_and_analyses/study_region/Study Region Extent.Rdata')) {
  load('./data_and_analyses/study_region/Study Region.Rdata')
  load('./data_and_analyses/study_region/Study Region Extent.Rdata')
} else {
  studyRegion <- rgdal::readOGR(dsn = './data_and_analyses/study_region', 'study_region')
  studyRegion <- crop(studyRegion, extent(-178.2166, 83.7759, -55.90223, 83.6236))
  projection(studyRegion) <- getCRS("WGS84")
  studyExtent <- extent(studyRegion)
  save(studyRegion, 
       file='./data_and_analyses/study_region/Study Region.Rdata', 
       compress=TRUE)
  save(studyExtent, 
       file='./data_and_analyses/study_region/Study Region Extent.Rdata', 
       compress=TRUE)
}

# clip environmental PCAs to study extent for given species, visualize, and save:
envDataClipped <- list()
for (i in 1:nlayers(envPresences)) {
  x <- envPresences[[i]]
  x <- crop(x, studyExtent)
  x <- mask(x, studyRegion)
  projection(x) <- getCRS("WGS84")
  envDataClipped[[i]] <- x
  envData <- stack(envDataClipped)
  # plot(envData)
}
```

Match environmental data to records.
```{r}
occsEnv <- raster::extract(envData, 
                              cbind(recordsRaw$longitude, 
                                    recordsRaw$latitude))
occsEnvDf <- as.data.frame(occsEnv)
records <- cbind(recordsRaw, occsEnvDf)
```

Remove records that fall in the water & visualize.
```{r}
# records in the water:
if (any(is.na(rowSums(occsEnvDf)))) {
  water <- records[which(is.na(rowSums(occsEnvDf))), ] 
  water <- SpatialPointsDataFrame(water[,ll], data = water,
                                proj4 = getCRS('wgs84', TRUE))
}

# remove records in water from dataset:
if (any(is.na(rowSums(occsEnvDf)))) records <-
  records[-which(is.na(rowSums(occsEnvDf))), ]

# visualize the points that fall in the water
# convert to sp object
recordsSp <- SpatialPointsDataFrame(records[, ll], data = records,
                                       proj4 = getCRS('wgs84', TRUE))

# visualize
plot(recordsSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' occurrences (BIEN) thinned'), list(x=sp)))
if (exists("water")) {
  plot(water, col = 'blue', add = TRUE)
}
map("state", add = TRUE)
map("world", add = TRUE)
```

Original buffer (110 km)
```{r}
# range_buffer_final = 110 km buffer
# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(range_buffer_final, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.110 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.110) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

160 km buffer
```{r}
# calculate 160 km buffer
library(units)
library(sf)
buffer_distance <- as_units(160, "km")
rangeMapSf <- st_transform(st_as_sf(x = range), getCRS('albersNA'))
range_buffer <- st_buffer(rangeMapSf, dist = buffer_distance)
buffer.160 <- st_union(range_buffer)

# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(buffer.160, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.160 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.160) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

320 km buffer
```{r, warning = FALSE}
# calculate 320 buffer
buffer_distance <- as_units(320, "km")
rangeMapSf <- st_transform(st_as_sf(x = range), getCRS('albersNA'))
range_buffer <- st_buffer(rangeMapSf, dist = buffer_distance)

# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(buffer.160, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.320 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.320) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), 'background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.110 <- raster::extract(climate, randomBgSites.110)
randomBgEnv.110 <- as.data.frame(randomBgEnv.110)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.110))
if (any(isNa)) {
    randomBgSites.110 <- randomBgSites.110[-which(isNa), ]
    randomBgEnv.110 <- randomBgEnv.110[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.110 <- randomBgSites.110[1:10000,]
randomBgEnv.110 <- randomBgEnv.110[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.110 <- cbind(randomBgSites.110, randomBgEnv.110)
names(randomBg.110)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.110, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.110)))

env <- rbind(occsEnv, randomBgEnv.110)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel4.110 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel4.110, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4')
# prediction
envMap4.110 <- predict(
  climate[[predictors]],
  envModel4.110,
  filename = paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer_0KYBP_pc4'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp4.110 <- rasterToPolygons(envMap4.110)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.110, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap4.110, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

***********************************
Using 160 km buffer:
```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.160 <- raster::extract(climate, randomBgSites.160)
randomBgEnv.160 <- as.data.frame(randomBgEnv.160)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.160))
if (any(isNa)) {
    randomBgSites.160 <- randomBgSites.160[-which(isNa), ]
    randomBgEnv.160 <- randomBgEnv.160[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.160 <- randomBgSites.160[1:10000,]
randomBgEnv.160 <- randomBgEnv.160[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.160 <- cbind(randomBgSites.160, randomBgEnv.160)
names(randomBg.160)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.160, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.160)))

env <- rbind(occsEnv, randomBgEnv.160)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel4.160 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel4.160, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4')
# prediction
envMap4.160 <- predict(
  climate[[predictors]],
  envModel4.160,
  filename=paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer0KYBP'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp4.160 <- rasterToPolygons(envMap4.160)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.160, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap4.160, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

***********************************
Using 320 km buffer:
```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.320 <- raster::extract(climate, randomBgSites.320)
randomBgEnv.320 <- as.data.frame(randomBgEnv.320)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.320))
if (any(isNa)) {
    randomBgSites.320 <- randomBgSites.320[-which(isNa), ]
    randomBgEnv.320 <- randomBgEnv.320[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.320 <- randomBgSites.320[1:10000,]
randomBgEnv.320 <- randomBgEnv.320[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.320 <- cbind(randomBgSites.320, randomBgEnv.320)
names(randomBg.320)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.320, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.320)))

env <- rbind(occsEnv, randomBgEnv.320)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel4.320 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel4.320, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4')
# prediction
envMap4.320 <- predict(
  climate[[predictors]],
  envModel4.320,
  filename=paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer0KYBP'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp4.320 <- rasterToPolygons(envMap4.320)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.320, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap4.320, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.110, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)

plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.160, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)

plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap4.320, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
```

---
title: "Finding perfect model"
author: "Lauren Jenkins"
date: "20 December 2021"
output: html_document
---

Just playing with buffer sizes & pca's

Setup environment:
```{r}
rm(list = ls())
library(dplyr)
library(ggplot2)
library(tidyr)

library(raster)
library(rgdal)
library(enmSdm)
library(terra)
library(maxnet)
library(sf)

library(maps)
library(rnaturalearth)
library(rnaturalearthdata)
library(rnaturalearthhires)

setwd('/Volumes/LJ MacBook Backup/MOBOT/PVMvsENM')
```

Load PCA environment and set constants:
```{r}
# load PCA data
load('./PCA_Beyer')

# set constants
ll <- c('longitude', 'latitude')
sp <- 'Fraxinus pennsylvanica'
species <- 'fraxinus_pennsylvanica'
speciesAb <- 'FraxPenn'
speciesAb_ <- 'Frax_Penn'
```

Load records for given species.
```{r}
recordsRaw <- paste0('./cleaning_records/', species, '_finalRecords.rData')
load(recordsRaw)
recordsRaw <- data.frame(speciesSf_filtered_final) # 542 observations
recordsRaw$geometry <- gsub("[c()]", "", recordsRaw$geometry)
recordsRaw <- separate(data = recordsRaw, 
                    col = 'geometry', 
                    into = ll, 
                    sep = "\\,")
recordsRaw$longitude <- as.double(recordsRaw$longitude)
recordsRaw$latitude <- as.double(recordsRaw$latitude)
```

Load cleaned records data.
```{r}
load(paste0('./03 - Modeling Workspace - ', speciesAb_, ' Cleaning'))
rangeMap <- littleRange_FraxPenn
```

Load environmental PCA rasters. If already clipped, load that. Otherwise, clip the data to present.
PC1 = 44.44%
PC2 = 71.67%
PC3 = 81.05%
PC4 = 87.65%
PC5 = 91.6%
PC6 = 96.39%
```{r}
pcPrediction <- list()
# label each env layer by variable and year
for (i in 1:length(clim)) {
  names(clim[[i]]) <- c("bio1", "bio10", "bio11", "bio12", "bio13", "bio14", "bio15", "bio16",
                        "bio17", "bio18", "bio19", "bio4", "bio5", "bio6", "bio7",
                        "bio8", "bio9", "cloudiness", "relative_humidity")
  pcPrediction[i] <- raster::predict(clim[[i]], pca, index = 1:5)
  names(pcPrediction[[i]]) <- paste0("pc", 1:5, "_", (i-1)*1000, "KYBP")
}

envDataPca <- stack(pcPrediction)

# keep only rasters (first three rasters) from present (0 KYBP)
envPresences <- stack(envDataPca[[1:5]])
names(envPresences) <- paste0('pca', 1:5)

# check projections = wgs84
print("Ensure that the projection of these rasters is WGS84:")
print(paste0("Projection of envPresences = ", projection(envPresences)))

# define study region & extent, load if already defined
if (file.exists('./data_and_analyses/study_region/Study Region.Rdata') & 
    file.exists('./data_and_analyses/study_region/Study Region Extent.Rdata')) {
  load('./data_and_analyses/study_region/Study Region.Rdata')
  load('./data_and_analyses/study_region/Study Region Extent.Rdata')
} else {
  studyRegion <- rgdal::readOGR(dsn = './data_and_analyses/study_region', 'study_region')
  studyRegion <- crop(studyRegion, extent(-178.2166, 83.7759, -55.90223, 83.6236))
  projection(studyRegion) <- getCRS("WGS84")
  studyExtent <- extent(studyRegion)
  save(studyRegion, 
       file='./data_and_analyses/study_region/Study Region.Rdata', 
       compress=TRUE)
  save(studyExtent, 
       file='./data_and_analyses/study_region/Study Region Extent.Rdata', 
       compress=TRUE)
}

# clip environmental PCAs to study extent for given species, visualize, and save:
envDataClipped <- list()
for (i in 1:nlayers(envPresences)) {
  x <- envPresences[[i]]
  x <- crop(x, studyExtent)
  x <- mask(x, studyRegion)
  projection(x) <- getCRS("WGS84")
  envDataClipped[[i]] <- x
  envData <- stack(envDataClipped)
  # plot(envData)
}
```

Match environmental data to records.
```{r}
occsEnv <- raster::extract(envData, 
                              cbind(recordsRaw$longitude, 
                                    recordsRaw$latitude))
occsEnvDf <- as.data.frame(occsEnv)
records <- cbind(recordsRaw, occsEnvDf)
```

Remove records that fall in the water & visualize.
```{r}
# records in the water:
if (any(is.na(rowSums(occsEnvDf)))) {
  water <- records[which(is.na(rowSums(occsEnvDf))), ] 
  water <- SpatialPointsDataFrame(water[,ll], data = water,
                                proj4 = getCRS('wgs84', TRUE))
}

# remove records in water from dataset:
if (any(is.na(rowSums(occsEnvDf)))) records <-
  records[-which(is.na(rowSums(occsEnvDf))), ]

# visualize the points that fall in the water
# convert to sp object
recordsSp <- SpatialPointsDataFrame(records[, ll], data = records,
                                       proj4 = getCRS('wgs84', TRUE))

# visualize
plot(recordsSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' occurrences (BIEN) thinned'), list(x=sp)))
if (exists("water")) {
  plot(water, col = 'blue', add = TRUE)
}
map("state", add = TRUE)
map("world", add = TRUE)
```

Original buffer (110 km)
```{r}
# range_buffer_final = 110 km buffer
# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(range_buffer_final, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.110 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.110) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

160 km buffer
```{r}
# calculate 160 km buffer
library(units)
library(sf)
buffer_distance <- as_units(160, "km")
rangeMapSf <- st_transform(st_as_sf(x = range), getCRS('albersNA'))
range_buffer <- st_buffer(rangeMapSf, dist = buffer_distance)
buffer.160 <- st_union(range_buffer)

# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(buffer.160, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.160 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.160) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), ' background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

320 km buffer
```{r, warning = FALSE}
# calculate 320 buffer
buffer_distance <- as_units(320, "km")
rangeMapSf <- st_transform(st_as_sf(x = range), getCRS('albersNA'))
range_buffer <- st_buffer(rangeMapSf, dist = buffer_distance)

# use the buffer/calibration region for extracting background sites
calibRegionSpAlb <- as(buffer.160, 'Spatial')
calibRegionSpAlb <- sp::spTransform(calibRegionSpAlb, getCRS('albersNA', TRUE))
calibRegionSpWgs <- sp::spTransform(calibRegionSpAlb, getCRS('wgs84', TRUE))

# get 20,000 random background sites from calibration region
bgTestSpAlb <- spsample(calibRegionSpAlb, n=20000, type='random')
bgTestSp <- sp::spTransform(bgTestSpAlb, getCRS('wgs84', TRUE))
randomBgSites.320 <- as.data.frame(coordinates(bgTestSp))
names(randomBgSites.320) <- ll

# sanity check: plot the background sites
plot(bgTestSp, pch = 16, cex = 0.5, col = "red", 
     main = substitute(paste(italic(x), 'background sites'), list(x=sp)))
plot(calibRegionSpWgs, add = TRUE, border = 'blue')
map("state", add = TRUE)
map("world", add = TRUE)
```

```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.110 <- raster::extract(climate, randomBgSites.110)
randomBgEnv.110 <- as.data.frame(randomBgEnv.110)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.110))
if (any(isNa)) {
    randomBgSites.110 <- randomBgSites.110[-which(isNa), ]
    randomBgEnv.110 <- randomBgEnv.110[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.110 <- randomBgSites.110[1:10000,]
randomBgEnv.110 <- randomBgEnv.110[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.110 <- cbind(randomBgSites.110, randomBgEnv.110)
names(randomBg.110)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.110, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.110)))

env <- rbind(occsEnv, randomBgEnv.110)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel5.110 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel5.110, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4', 'pca5')
# prediction
envMap5.110 <- predict(
  climate[[predictors]],
  envModel5.110,
  filename = paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer_0KYBP_pc4'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp5.110 <- rasterToPolygons(envMap5.110)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.110, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap5.110, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

***********************************
Using 160 km buffer:
```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.160 <- raster::extract(climate, randomBgSites.160)
randomBgEnv.160 <- as.data.frame(randomBgEnv.160)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.160))
if (any(isNa)) {
    randomBgSites.160 <- randomBgSites.160[-which(isNa), ]
    randomBgEnv.160 <- randomBgEnv.160[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.160 <- randomBgSites.160[1:10000,]
randomBgEnv.160 <- randomBgEnv.160[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.160 <- cbind(randomBgSites.160, randomBgEnv.160)
names(randomBg.160)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.160, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.160)))

env <- rbind(occsEnv, randomBgEnv.160)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel5.160 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel5.160, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4')
# prediction
envMap5.160 <- predict(
  climate[[predictors]],
  envModel5.160,
  filename=paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer0KYBP'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp5.160 <- rasterToPolygons(envMap5.160)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.160, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap5.160, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

***********************************
Using 320 km buffer:
```{r}
# extract environment at background sites
climate <- envData
randomBgEnv.320 <- raster::extract(climate, randomBgSites.320)
randomBgEnv.320 <- as.data.frame(randomBgEnv.320)

# remove any sites with NA for at least one variable
isNa <- is.na(rowSums(randomBgEnv.320))
if (any(isNa)) {
    randomBgSites.320 <- randomBgSites.320[-which(isNa), ]
    randomBgEnv.320 <- randomBgEnv.320[-which(isNa), ]
}

# only keep 10,000 random background sites
randomBgSites.320 <- randomBgSites.320[1:10000,]
randomBgEnv.320 <- randomBgEnv.320[1:10000,]
# combine with coordinates and rename coordinate fields
randomBg.320 <- cbind(randomBgSites.320, randomBgEnv.320)
names(randomBg.320)[1:2] <- ll

# dir.create('./Background Sites', recursive=TRUE, showWarnings=FALSE)
# save(randomBg.320, 
#      file = paste0('./Background Sites/Random Background Sites across Study Region - ', 
#                    speciesAb, '.Rdata'),
#     compress=TRUE)
```

Ready for maxent!
Turn off clamping to avoid extrapolating (in maxnet(), clamp = FALSE).
Also, type = cloglog (to make the output between 0 & 1).
```{r}
presBg <- rep(c(1, 0), c(nrow(recordsRaw), nrow(randomBg.320)))

env <- rbind(occsEnv, randomBgEnv.320)
env <- cbind(presBg, env)
env <- as.data.frame(env)
env <- na.omit(env)

# create output directory for model object and rasters
dir.create(paste0('./Models/', speciesAb_, '_Maxent'),
    recursive=TRUE, showWarnings=FALSE)

# model species
envModel5.320 <- enmSdm::trainMaxNet(data = env, resp = "presBg")
# envModel <- maxnet(p = presBg, data = trainData)
modelFileName <- paste0('./Models/', speciesAb_, '_Maxent/Model.Rdata')
save(envModel5.320, file = modelFileName, compress=TRUE)

predictors <- c('pca1', 'pca2', 'pca3', 'pca4', 'pca5')
# prediction
envMap5.320 <- predict(
  climate[[predictors]],
  envModel5.320,
  filename=paste0('./Models/', speciesAb_, '_Maxent/maxentPredictionBeyer0KYBP'), 
  clamp = F, 
  format='GTiff', 
  overwrite = TRUE, 
  type='cloglog')

envMapSp5.320 <- rasterToPolygons(envMap5.320)
```

Visualize model output, focused on species.
```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.320, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
map("state", add = TRUE)
map("world", add = TRUE)
points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

Visualize output, zoomed out to study region.
```{r}
plot(envMap5.320, main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(rangeMap, border = 'blue', add = TRUE)
# map("state", add = TRUE)
# map("world", add = TRUE)
# points(records$longitude, records$latitude, pch = 16, cex = 0.6, col = 'red')
```

```{r}
plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.110, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)

plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.160, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)

plot(rangeMap, border = 'blue', main = substitute(paste('Maxent output, ', 
                                                        italic(x),
                                                        ' occurrences'), list(x=sp)))
plot(envMap5.320, add = TRUE)
plot(rangeMap, border = 'blue', add = TRUE)
```

